Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-08 robdml.tex
tex

% inspiration.
Recall that a DML model is already prone to collapse with very hard but not
adversarially purturbed sample triplets~\cite{facenet}.
%
To overcome this issue, various triplet sampling strategies, such as
semi-hard~\cite{facenet} and distance~\cite{distance} are proposed to provide
effective negative examples to facilitate metric learning while avoiding model
collapse.
%
In the classical min-max adversarial training, the adversarial example triplet
created by maximizing the triplet loss value can incur model collapse more
easily than unperturbed sample triplets.
%

In this paper, we propose Hardness Manipulation (HM) to create adversarial
examples for increasing their hardness towards that of a harder sample triplet.
%
%
%
Triplet hardness still matters in adversarial training for deep metric
learning.

>>>>

% inspiration.
Recall that a DML model is already prone to collapse with very hard but not
adversarially purturbed sample triplets~\cite{facenet}.
%
To overcome this issue, various triplet sampling strategies, such as
semi-hard~\cite{facenet} and distance~\cite{distance} are proposed to provide
effective negative examples to facilitate metric learning while avoiding model
collapse.
%
In the typical min-max adversarial training~\cite{madry}, the adversarial
example triplet created by maximizing the triplet loss value can incur model
collapse more easily than unperturbed sample triplets.
%
But what if we do not maximize the triplet loss value to the theoretical upper
bound, but merely increase it towards a value that could be caused by
a known, unperturbed and harder sample triplet?

% HM
In this paper, we propose Hardness Manipulation (HM) to make adversarially
training a DML model with triplet loss feasible without model collapse,
so the model could efficiently learn from a strong adversary.
%
First, we sample two triplets, namely ``source triplet'' and ``destination
triplet'' with the same anchor example but in two different pre-defined
sampling strategy.
%
The ``destination triplet'' is expected to result in a higher loss value.
%

create adversarial
examples for increasing their hardness towards that of a harder sample triplet.
%
%
%
Triplet hardness still matters in adversarial training for deep metric
learning.
